The Gaia-ESO Survey: Combined estimates from multiple stellar analysis pipelines

### 2016-11-29 article/ms.tex
Add words about flag propagation

% Stellar parameter homogenisation
\subsection{Stellar parameter homogenisation}
\label{sec:stellar-parameter-homogenisation}

Flag propagation.

\noindent{}We make the following assumptions when constructing a model for homogenisation:

@@

We list the nomenclature and descriptions for all model parameters in Table 
\ref{tab:model-parameters}. The number of total model parameters is given by
\begin{eqnarray}
    N_{parameters} = N_{calibrators} + N_{missing} + 6N_{nodes} + \frac{N_{nodes}!}{2(N_{nodes} - 2)!}
\end{eqnarray}
where $N_{calibrators}$ accounts for the true values of the calibrators, the
$N_{missing}$ models the missing data points for the calibrator spectra, the
$6N_{nodes}$ summarises parameters that model biases, as well as systematic
and random uncertainties. The final term $\frac{N_{nodes}!}{2(N_{nodes} - 2)!}$
is the number of correlation coefficients between all nodes.\footnote{
This inference problem is numerically unstable, as there are many model parameters
which are used toi

 construct a covariance matrix that is sampled

In practice
we include the Cholesky decomposition $L$ of the $N \times N$ correlation matrix $\rho$
as model parameters instead of directly modeling the $\rho$ terms.  The correlation
matrix can be reconstructed by $\rho = (IL)\cdot(IL)^T$ where $I$ is the $N \times N$
identity matrix. This parameterisation helps ensure that the resulting covariance
matrix $\mathbf{\Sigma}$ is positive semi-definite, and can therefore be used for
sampling.}

@@

>>>>

% Stellar parameter homogenisation

\subsection{Flag propagation}
\label{sec:flag-propagation}

In earlier data releases the analysis nodes were \emph{encouraged} to submit flags that
describe technicalities that may have affected their analysis, or peculiar issues
that were identified in a star or spectrum.  These flags proved invaluable in enabling 
homogenisation leads to identify scenarios where particular pipelines perform poorly. 
For example, OACT (WG11) were able to identify and account for high rotation, whereas 
ULB (WG11) would not identify fast rotators, and ULB (WG11) would report significantly
noisier stellar parameters for fast rotators.

From DR5 onwards, these examples led to the decision to \emph{require} analysis nodes 
to submit stellar parameters ($\teff$, $\logg$, [Fe/H], and $\xi$ where appropriate),
for every spectrum, or supply a technical flag explaining why no stellar parameters
could be derived (preferably supply both when appropriate). The full list of flags used
in DR5 is listed in Appendix \ref{appendix:flags}.  In Figures \ref{fig:flag-propagation-wg10}
and \ref{fig:flag-propagation-wg11} we show heatmaps of flag usage between different 
analysis nodes in WG10 and WG11, respectively.  If all nodes reported the same flags then
a one-to-one density mapping would be visible.  Note that here we are only showing instances
where multiple flags have been reported for a single spectrum: for the sake of clarity we
do not show situations where only one flag is reported by a single node.  Figures 
\ref{fig:flag-propagation-wg10} and \ref{fig:flag-propagation-wg11} demonstrate that when one analysis node reports a flag, it
is highly unlikely that the same issue has been reported by another node. The most common
situation is that one node will report a descriptive flag (e.g., \texttt{10106} 
to indicate a broken spectrum containing a picket-fence or Heaviside pattern), and other
nodes will report a non-specific flag (e.g., \texttt{10302}: code convergence issue).

We constructed a sequence of rules to propagate detailed flag information to other results.
We categorised these rules to propagate to results that were derived from either the same 
spectrum (e.g., \texttt{10106} broken spectrum), or the same star (e.g., suspected binary 
star system).  These propagation rules are listed in Tables \ref{tab:flag-propagation-spectrum}
and \ref{tab:flag-propagation-cname}. Any results with propagated technical flags were 
marked as failing internally quality control standards, and were not used in the homogenisation
process.  Results with any of the flags listed in Table \ref{tab:node-specific-flags} were
were also excluded from homogenisation, however those flags were considered `node-specific'
in that they do not necessarily require propagation to other results derived from the same
spectrum or star.

We also excluded all cluster records (\texttt{GES\_TYPE} of 'XX-XX-XX' or `XX-XX-XX') 
reported by OACT (WG11).  Although OACT performed excellently in the benchmark and blind tests
in WG11, we identified convergence issues in cluster field, which consequently altered the
metallicity scale of globular clusters.

% Any other exclusions? WG10 IAC results at edges?

The total number of records (set of $\teff$, $\logg$, [Fe/H]) that were marked as failing
internal quality controls due to reported or propagated flags were XXX for WG10 and XXX for
WG11.  This represents a small fraction of the total number of records submitted in each
working group (XXX\% for WG10 and XXX\% for WG11).  The number of stars (unique \texttt{CNAME}s) where all results were excluded due to reported or propagated flags was XXX for WG10 and XXX for WG11.  These cases were principally due to data reduction issues (e.g., flags 
\texttt{10100-10150}), and these examples were supplied to the data reduction group to
examine before the sixth data release.

\subsection{Stellar parameter homogenisation}
\label{sec:stellar-parameter-homogenisation}

\noindent{}We make the following assumptions in our homogenisation model:

@@

We list the nomenclature and descriptions for all model parameters in Table 
\ref{tab:model-parameters}. The number of total model parameters is given by
\begin{eqnarray}
    N_{parameters} = N_{calibrators} + N_{missing} + 6N_{nodes} + \frac{N_{nodes}!}{2(N_{nodes} - 2)!}
\end{eqnarray}
where $N_{calibrators}$ accounts for the true values of the calibrators, the
$N_{missing}$ models the missing data points for the calibrator spectra, the
$6N_{nodes}$ summarises parameters that model biases, as well as systematic
and random uncertainties. The final term $\frac{N_{nodes}!}{2(N_{nodes} - 2)!}$
is the number of correlation coefficients between all nodes.\footnote{
This inference problem is numerically unstable because a \emph{positive semi-definite}
covariance matrix must be constructed from the model parameters in every MCMC draw, 
and there is no joint prior available to ensure that the positive semi-definite 
condition is met.  Although there are parameterisations that will help enforce 
this condition \citep{Pinheiro:1996}, in practice we found that including 
the Cholesky decomposition $L$ of the $N \times N$ correlation matrix $K$ as model
parameters --- instead of directly modelling the set of node correlation 
coefficients $\{\rho\}$ --- was sufficient to ensure positive-semidefinite 
covariance matrices and produce sensible solutions.  The correlation matrix can 
then be reconstructed by $K = (IL)\cdot(IL)^T$ where $I$ is the $N \times N$ identity
matrix.  This representation implies a uniform prior on all correlation coefficients
$p\left(\rho\right) = \mathcal{U}\left(-1, +1\right)$.}

@@

\begin{thebibliography}{}
\bibitem[Pinheiro \& Bates(1996)]{Pinheiro:1996} Pinheiro, J.~C., \& Bates, D.~M.\ 1996, Statistics and Computing, 6, 3

\end{thebibliography}
